Physics-based protein structure prediction and design using the confinement method

### 2012-10-19 Paper/paper.tex
third draft arijit

In computational structural biology there are numerous cases where free energy between well defined
states are necessary. Examples ranges from small to large conformational change of protein due to
ligand binding, change of pH etc. Free energy also play an important role in case of protein
folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
structures of small globular proteins have a unique, thermodynamically stable native structure with
their conformation at the global free energy minimum. Regardless of the starting point or unfolding
it by changing different condition most proteins will finally assume the same structure. It is very
important for proteins to achieve their native conformation since defects in protein folding may be
the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
prion when enters any healthy organism it converts properly folded protein into misfolded one.
Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
the protein. These observations further give free energy a special importance in structural biology.

However, the theoretical calculation of conformational free energy change become difficult if the
states of interest are very different. In such scenario timescale involved in such tranformation may
be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
Method like umbrella sampling, thermodynamic integration along with classical MD can suffer from
overlap problem and require large computational effort. The idea of reaction coordinate became more
complicated in case of protein folding.  As free energy landscape of the protein folding is rough \cite{Dill1997},
it can be trapped in a energy well during its visit in the conformational space. Even with the
advanced method like replica exchange molecular dynamics the three dimensional structure may be
trapped in a local minima for a considerable time. Due to various reasons it can also be misfolded
and trapped in a local energy minima. And one can mistakenly identify the misfolded state as the
native state (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).

Calculation of free energy has been successfully attempted by a number of groups \cite{Ytreberg2006} - \cite{Zheng2008}. In
recent years methods like Refrence system method, Decativated Morphing \cite{Park2008}, Orthogonal Space Random
Work, Confinement Method have came out for calculation of free energy method. Some of the groups
relies on one of the great strength of free energy calculation, that it is a state function and thus
it is independent of path.

In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
Cecchini et. al. \cite{Cecchini2009}. This method relies on a thermodynamic cycle where first, non-harmonic degrees of
freedom are removed by applying a series of restraint to the system. Then the free energy of the
remaining harmonic system is obtained using a normal mode calculation and combined with the
non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
first reproduced that known result. Then we have applied this method to larger proteins for the
first time. The free energy of conformation of pair of proteins with similar sequence but completely
different fold was calculated. Encouraged by the result, we applied the method of different targets
of CASP9 competition.

@@

\Justin{Arijit, Al: did we do any other control experiments on small peptides that we should report
here?}

Previous applications of the confiment method have focused on relatively short peptides, up to $17$
residues in length. In this work we apply the confinement method to
larger proteins. One of the  applications we explore in the present work is the re-scoring, or
metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
(CASP) experiment (described in detail later). We computed the relative free energies of predictions
submitted during CASP9 10, with the expectation that the most native-like prediction will have
the lowest free energy. As a control, for several targets we also calculated the relative free
energy of the experimental structure (once it was available from the PDB). As
Table~\ref{table:casp_control} shows, in X of Y cases, the confinement method correctly assigns a
lower free energy to the experimentally determined structure than to any of the decoys. Although
differentiating between experimentally determined structures and computer generated predictions is
not a stringent test of a scoring method---for example, examining the residue-residue packing can
easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
"sanity check".

@@

\begin{center}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $7.70 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline 
\end{tabular}
\end{center}

@@

The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
identifies the correct structure for all six sequences. The free energy differences range from
around 2.9 to 5.7 kcal/mol, which is consistent with small changes in sequences and with estimates
made by Orban and co-workers \cite{He2008}, \cite{{Alexander2009}}.

\begin{figure}
\includegraphics[width=14cm,height=14cm]{orban.pdf}
\label{fig:orban}
\caption{The confinement method correctly predicts the structural preferences of six chameleon
sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
the two folds are reported for each sequence.}
\end{figure}

%\begin{table}
%\caption{The confinement method correctly reproduces the structural preferences of a series of
%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
%helix bundle ($3 \alpha$).}
%\label{table:orban}
%\begin{center}
%\begin{tabular}{l l l}\hline
%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
 %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
%GA98       & \Justin{Arijit: fill in} &                           \\ \hline
%GB98       &                          &                           \\ \hline
%\end{tabular}
%\end{center}
%\end{table}

\subsection{The confinement method correctly identifies the most native-like predictions from a
subset of CASP predictions}

Next, we have applied the confinement method to several targets from the CASP9 and CASP10
experiments. CASP is a world-wide competition aiming at finding the state of the art methodologies
in protein folding. Groups, both from knowledge based methods and physics based methods participate
in this blind competition and for each sequence target they can submit five models. The predictors
were also asked to calculate their best model among the set of five they submitted. However, in many
cases predictors were unable to predict/rank their models. Using the confinement method, we aim to
rank protein structures using free energy as a scale. In recent CASP experiments Global distance
test scose (GDT\_TS) method is used for final analysis of different submitted structures with the
NMR/crystallographically solved protein native structure. GDT\_TS represent the average percentage
of residues that are in close proximity in two structures optimally superimposed using four
different distance cutoff. Although GDT\_TS alone is not a reliable measure for the difficult
modeling cases, it has been well accepted method in recent years among the community. We have also
compared our free energy result with the GDT\_TS score for different cases we studied.

%The GDTTS value of a model is determined as follows. A large sample of possible structure
%superpositions of the model on the corresponding experimental structure is generated by superposing
%all sets of three, five, and seven consecutive C$\alpha$ atoms along the backbone (each peptide
%segment provides one superposition). Each of these initial superpositions is iteratively extended,
%including all residue pairs under a specified threshold in the next iteration, and continuing until
%there is no change in included residues. The procedure is carried out using thresholds of 1, 2, 4,
%and 8 $\AA$, and the final superposition that includes the maximum number of residues is selected
%for each threshold.

\subsubsection{T0559}

The native structure is a  67 residue NMR solved structure. We first reviewed all the five models
submitted by the group "BAKER-ROSETTASERVER", which was the best predictor group for this particular
target.  Among them we choose only the three structures, model 1, 3 and 5 to calculate the free
energy and subsequently identify the most stable structure among them. The rest of the two
structures are close to either of the selected structure and thus they are ommited from our
calculation. According to the GDT\_TS score, the model 1 was picked up correctly from the set of
structure generated . However, the order of model 3 and 5 was not correct.
It will be interesting to know whether confinement method can order the strcutures in terms
of free energy scale. The result is presented in Figure~\ref{fig:T0559}.

\subsubsection{T0560}

Target T0560 is a 74 amino acid residue target from CASP9. However residues 3-66 were kept for final
result.  For this target we have compared the free energy of the model 2 and model 4 submitted by
the group "Splicer" along with the crystal structure.  The main difference between the model 2 and 4
is the orientation of the ....  The final result is presented in Figure~\ref{fig:T0560}.

\subsubsection{T0538}

Target T0538 was a 54 residue target. The native structure was generated using NMR. The best model
that is closest to the native structure was generated by the group PconsR. However, their model 3
was better than their model 1. This structure was quite close to the NMR structure and had GDT\_TS
value of 96.23. We have calculated the free energy between this structure and the best structure
submitted by the group "Shell", "FOLDIT" and "BAKER-ROSETTASERVER". The calculated free energy value
is presented in Figure~\ref{fig:T0538}.

\subsubsection{T0540}

Target T0540 was a 90 residue target. We have taken the best models from the group "LTB" which was
the best performing group for this target and also from the group Mufold and calculated the free
energy difference between these models with that of the native crystal structure.  As expected, the
crystal structure has the lowest free energy value, followed by Model 1 with GDT\_TS score of
$69.72$ and model 2 with GDT\_TS value of $53.789$. The final result is presented in
Figure~\ref{fig:T0540}.

\subsubsection{T0531}

This is a 65 amino acid residue target with residues 6-63 were kept for final analysis. We have
calculated the free energy difference between all the five models submitted by the group
"MUFOLD-MD", which was one of the best performed group for this target.  The result is presented in
Figure~\ref{fig:T0531}.

\subsubsection{T0605}

This is a 72 amino acid residue protein but for the final calculation only the residues 18-66 amino acids were kept for the final analysis.

The best performing group was "Baker" for this target. We have carried out our calculation with the model first, second and fifth for the
final calculation. The rest two model had closer GDT\_TS value compared to the previously mentioned models and thus they were omitted from
furthe calculation. The result is presented in Figure~\ref{fig:T0605}. (This is a case where we can demand the confinement
method is better than GDT\_TS).

\begin{figure}
\begin{center}
\includegraphics[width=14cm,height=14cm]{method.pdf}
\end{center}
\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
\label{fig:method}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{bhp.pdf}
\end{center}
\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
\label{fig:bhp_conf}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{G95.pdf}
\end{center}
\caption{Two protein with 95 \% similar sequence but different folds}
\label{fig:G95}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0559.pdf}
\end{center}
\caption{Target T0559 from CASP9}
\label{fig:T0559}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0538.pdf}
\end{center}
\caption{Target T0538}
\label{fig:T0538}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0560.pdf}
\end{center}
\caption{Target T0560}
\label{fig:T0560}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0540.pdf}
\end{center}
\caption{T0540}
\label{fig:T0540}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0531.pdf}
\end{center}
\caption{T0531}
\label{fig:T0531}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0605.pdf}
\end{center}
\caption{T0605}
\label{fig:T0605}
\end{figure}

@@

@@

>>>>

In computational structural biology there are numerous cases where free energy between well defined
states are necessary. Examples ranges from small to large conformational change of protein due to
ligand binding, change of pH etc \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
Free energy also play an important role in case of protein
folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
structures of small globular proteins have a unique, thermodynamically stable native structure with
their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting point or unfolding
it by changing different condition most proteins will finally assume the same structure. It is very
important for proteins to achieve their native conformation since defects in protein folding may be
the molecular cause of a range of human genetic disorder. For example a misfolded protein known as
prion when enters any healthy organism it converts properly folded protein into misfolded one.
Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
the protein. These observations further give free energy a special importance in structural biology.

However, the theoretical calculation of conformational free energy change become difficult if the
states of interest are very different \cite{Meirovitch2007}. In such scenario timescale involved in such tranformation may
be beyond timescale of direct classical molecular dynamics simulation.  Generally, in order to
calculate free energy a pathway or reaction coordinate is created with the intermediate structures.
Method like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
overlap problem and require large computational effort. The idea of reaction coordinate became more
complicated in case of protein folding.  As free energy landscape of the protein folding 
is rough \cite{Dill1997}, \cite{Dill2008}
it can be trapped in a energy well during its visit in the conformational space. Even with the
advanced method like replica exchange molecular dynamics the three dimensional structure may be
trapped in a local minima for a considerable time. Due to various reasons it can also be misfolded
and trapped in a local energy minima. And one can mistakenly identify the misfolded state as the
native state (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).

Calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
\cite{Ytreberg2006} - \cite{Zheng2008}, . In recent years methods like Refrence system method \cite{Ytreberg2006}, 
Decativated Morphing \cite{Park2008}, Orthogonal Space Random 
Work, Confinement Method have came out for calculation of free energy method.

Some of the groups
relies on one of the great strength of free energy calculation, that it is a state function and thus
it is independent of path.

In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
Cecchini et. al. \cite{Cecchini2009}. The free energy is a state function and thus it can be independent of path. 
Confinement method relies on this fact and uses a thermodynamic cycle where first, non-harmonic degrees of
freedom are removed by applying a series of restraint to the system. Then the free energy of the
remaining harmonic system is obtained using a normal mode calculation and combined with the
non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
first reproduced that known result. Then we have applied this method to larger proteins for the
first time. The free energy of conformation of pair of proteins with similar sequence but completely
different fold was calculated. Encouraged by the result, we applied the method of different targets
of CASP9 competition.

@@

Previous applications of the confiment method have focused on relatively short peptides, up to $17$
residues in length. In this work we apply the confinement method to
larger proteins. On that direction, We first test the confinement method for a chamelon sequence and 
found that it can correctly pick out the prefential structure of the chamelon sequence. Another of the 
applications we explore in the present work is the re-scoring, or
metaprediction, of structures submitted during the Critical Assessment of Structure Prediction
(CASP) experiment (described in detail later). We computed the relative free energies of predictions
submitted during CASP9 10, with the expectation that the most native-like prediction will have
the lowest free energy. As a control, for several targets we also calculated the relative free
energy of the experimental structure (once it was available from the PDB). As
Table~\ref{table:casp_control} shows, in most cases, the confinement method correctly assigns a
lower free energy to the experimentally determined structure than to any of the decoys. 
%Although differentiating between experimentally determined structures and computer generated predictions is
%not a stringent test of a scoring method---for example, examining the residue-residue packing can
%easily distinguish between the two [ref Rosetta Holes]---the results none the less serve as a useful
%"sanity check".

@@

\begin{center}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline 
\end{tabular}
\end{center}

@@

The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
identifies the correct structure for all five sequences. One of the hypothesis for fold switching
of chamelon sequence is that the structural transitions require states with diminished stability.
It is widely believed that if the free energy of the native state and the alternative state is within
 a range of $-5 Kcal/Mol$ then  it can quickly change fold when the stability of the native state 
decreases. The stability of the native state can decrease for a number reason ranging from chemical 
modification, breaking of disulphide bond or mutation as in this case. The free energy differences, that
came out from our calculation ranges from
around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.

\begin{figure}
\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
\label{fig:orban}
\caption{The confinement method correctly predicts the structural preferences of six chameleon
sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
the two folds are reported for each sequence.}
\end{figure}

%\begin{table}
%\caption{The confinement method correctly reproduces the structural preferences of a series of
%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
%helix bundle ($3 \alpha$).}
%\label{table:orban}
%\begin{center}
%\begin{tabular}{l l l}\hline
%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
 %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
%GA98       &  &                           \\ \hline
%GB98       &                          &                           \\ \hline
%\end{tabular}
%\end{center}
%\end{table}

\subsection{The confinement method correctly identifies the most native-like predictions from a
subset of CASP predictions}
\Alberto{Arijit: we need a global table of success, rather than individual tables. The simplest form
    would have 3 columns: name of the PDB, number of structures tested, percent of success of
    picking 1st model}

This method is particularly well suited to pick out the native/native like structure from misfolded or decoys. We have
tested this using different models from CASP (critical assessment of structure
prediction). CASP is a blind test world wide competition in which different groups apply their
methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
structures (ranked from better to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
competition we observed no correlation at all between the ordering of structures submited by the
groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}}. The consequences of this go beyond
those five structures; the deeper meaning is that groups producing ensembles of structures are
generating structures that are better than the ones they submit, but they do not know about it. In
fact, when querrying different groups after the results were known, most groups agree that this is
the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
models in many different environments, from structures for drug design leads to designing more
stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
accuracy is the fact that ranking is often done via a potential energy function, which in many cases
lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
order some structures from the previous CASP9 experiment. Our second test was to participate as a
metapredictor group in CASP10. In both methods, our ranking is determined as a free energy measure,
and it is compared with the ranking given by a geometrical comparison between native and submitted
models called GDT\_TS (Global distance test score \cite{Zemla2003}. 
GDT\_TS represent the average percentage of residues that are in close proximity in two
structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).

We have chosen the inital models on which to test the methods based on server groups that have
traditionally done good in past CASP events. There is a high correlation between the structures that
are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
50, it is difficult for us to say anything about the models. Our approach is simple, with every
subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
the structures. In all cases, we have calculated the free energy with respect to the native structure either obtained from NMR or 
x-ray crystallography but we do not need to make comparisons to native in order to do rank ordering.

\subsubsection{CASP9}

The question that we want to answer here is whether we can distinguish between native structure and the submitted models? 
%Identifying the native state from a set of models is a very easy test, that most metapredictors
%already handle correctly, but it is also a first proof of principle. 
First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
The best predictor group for this target was "BAKER-ROSETTASERVER".
We initlally checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
whereas the order of model 3 and 5 was wrong. The main difference
between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
structure from the submitted models , it can also correctly rank all the models.

To further test the method we have taken the example of protein BT2368 from Bacteroides thetaiotaomicron. The PDB id of this 
74 amino acid residue protein is 2L02 and CASP target code is T0560. We have compared the free energy difference between 
the native structure and the two of the five submitted models from the group "Splicer". The remaining three models are discarded as 
again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
Figure~\ref{fig:T0560} and published in final CASP result is only for models with residues 3 to 66. In order to keep
consistency, we have also do our analysis with models and crystal structure consisting residues 3 to 66. As expected,
the native state is identified correctly, and the two other models is ranked in the correct order.

Our next test was between models from the best submissions of different groups. 
For this purpose we have chosen the x-ray crystallographic structure of 
fas apoptosis inhibitory protein molecule whose pdb id and CASP target codes are 3MX7 and T0540 respectively. 
This protein contain 90 amino acid residues with 8 beta strands. We have chosen best models from groups 
"LTB" and "Mufold" for our analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}. 
As presented in  Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches 
with the GDT\_TS score.

There are however some instances where the method can fail, specially when GDT scores are
very close. We have found this for an engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538).
This protein contains 54 amino acid residues with three alpha helixes and two very short beta strand. 
The model that is closest to the native structure was generated by the group PconsR with a GDT\_TS value of 96.23.
We have also considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our analysis along with
the native crystal struture. The models from PconsR, Shell and FOLDIT are labelled as model 1, model 2 and model 3 
respectively in Figure~\ref{fig:T0538}. As presented from Figure~\ref{fig:T0538}, one can found that the 
model from PconsR has lowest free energy value, making it a more stable structure. Although ranking of the other models
remain consistent, it may happen that the theoretical model produced by the group "PconsR" much more stable than the 
crystal structure. This is particularly difficult target for rank ordering as all the models those are considered for this 
calculation are within RMSD values of 2.6 \AA.

\subsubsection{What can we say about low resolution models?}

We mentioned earlier that our predictive abilities are greatly decreased with low model qualities. It is
better to have atleast one good model to pick out the native/native like structures from the rest of the decoy. 
In oorder to test this hypothesis we performed another test with target T0531 of CASP9, in which we compared the crystal 
structure to five models from the MUFOLD server. The pdb id of this extracellular domain of the jumping translocation 
breakpoint protein is 2KJX and the CASP target code is T0531.
Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly
identified as expected and 2.) Surprisingly there is a high level of correlation between the GDT
ranking and the free energy ranking, only two structures with GDT\_TS score less than  35  are ordered
incorrectly. It is worth to note that models 2 and 3 have the same GDT and very different free
energies, meaning that the actual ordering could change a lot \cite{Perez2012}.

\subsubsection{Difficulty in CASP Experiments}

It is often difficult to model CASP targets using confinement and release method. There are many cases,
where the sequence of a particular target are trimmed during actual result presentation. This is illustrated
with the crystal structure of the leucine zipper domain of cGMP dependent protein kinase I (pdb id: 3NMD
and CASP9 code: T0605) as shown in Figure~\ref{fig:T0605}. The actual CASP 9 target consisted of 72 amino acid residues, but
as only 18-66 amino acid residues could be detected in actual crystal structure of this protein, the final result was trimmed 
with respect to that crystal structure with 49 amino acid residues which consist of a single alpha helix. The amino acid residues 18 and 66 are shown in
three models submitted by the group "Baker". It is important to note that blind prediction of model ranking on the basis of free energy can be hugely 
affected in such scenario as the only major difference between those models are the trimmed parts, specially the region consisting residues 1-18.

\subsubsection{Can confinement method perform better quality assessment in protein structure prediction ?}
A part of CASP experiment is dedicated to the assessment of the quality of predicted model \cite{Kryshtafovych2011}. 
Most of the top performing group in this catagory use consensus approaches for quality assessment \cite{Wang2011}.
A review of CASP9 quality assessment also pointed out that the method which are based on clustering technique perform
much better compare to those methos which are based on analysis of individual model \cite{Kryshtafovych2011}. 
In this background we try to answer, whether confinement method can do better quality assessment compare to the best performing
group (MUFOLD-WQA) in this catagory.
We have again picked up the CASP taget T0538 to answer this question. We have chosen the Model 3 submitted by PconsR (GDT\_TS = 96) 
and model 5 from the MULTICOM-NOVEL ( GDT\_TS = 83). Both are server predicted model and they were further used for quality assessment 
by the group MUFOLD-WQA \cite{Wang2011}. It is important to note that the model from PconsR was most accurate predicted model for this 
target and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of MUFOLD-WQA. 
Our free energy analysis we have shown that model from Pconsr is better than the model from 
MULTICOM-NOVEL $(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -8.2 Kcal/Mol)$, which   
matches well with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order. 
There is no doubt that the consensus approaches can predict the model quality in a faster manner. But we expect that confinement
approach can predict the model quality in a relatively expensive but much accurate way.

Further Analysis is going on Target 560.

\subsubsection{CASP10}

We participated in CASP10 as a metapredictor group. Given the tigh schedule of the CASP competition,
we only attempted proteins under 100 residues in length. Our main problem during this part of the
competition. was that proteins were trimmed from the actual calculation.
%\Alberto{Justin,Arijit: Probably should not comment on metapredicting since we were not able to do
%    good due to the trimming factor. Instead, we should move forward with Justin's idea of using
%    structures that other meta predictors did badly at. What stage is that in?}

@@

Proteins are essential parts of organisms and participate in virtually every process within living cells. 
In order to understand different functions of proteins, detect diseases caused by them and design more efficient drugs 
it is often useful to have the three dimensional structure of the protein. Experimental studies using NMR and X-ray crystallography is 
not enough to detect large number of structures. In recent past there are number of studies directed towards understanding protein 
structure using theoretical tools. Experimentally, it is relatively easy to identify the native structure from the data obtained 
from x-ray crystallography or NMR. On the other hand using theoretical tools it is often difficult to identify the native structure 
using both physics based methods and knowledge based method. In order to rank structure from the available data, some groups use 
clustering algorithms and the structure of the
most populated cluster is picked out as the most probable structure. However, it may happen that the structure may be trapped and 
spend most of its time in a kinetic trap, which may give a wrong interpretation of the clustering result. On the other hand, some groups
also depend on potential energy function to pick out their best structure. In this paper, we have used a free energy method to identify
the native/native like structure. First, we have tested this method for a chameleon sequence, which has a $95$ percent sequence identity
but one of them has a $3 \alpha$ structure and another one has $4 \beta + \alpha$ structure. Using the current method, we can found that 
the native structure always has most stable structure. We have also tested this method for different targets of CASP experiments. We found
that in most cases, it can pick out the native/native like structure in terms of lowest free enrgy value.

However, there may be some limitation of this method in terms of size of the protein molecule. The normal mode analysis may not be
accurate for proteins larger than 100 amino acid residues. This can be approximated by removing the hydrogen molecules from 
the system. 
The method can be also give wrong result if all the models in the decoy are very poor. It may also be difficult to rank structures if 
all the structures are very very close to the native structure. But that does not matter at all as our main aim is to differentiate
between the good and the bad structure. In this article we have extensively tested this method using examples from CASP experiment.
This method can be computationally expensive in actual CASP experiment if one want rank thousands of submitted models in CASP 
experiments. However, in real life experiment with a particular target this method can give accurate result. The important 
fact about Confinement method is that all the part of the calculation can be done independently. We have mostly carried out 
our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we found that it take only 4 hour to complete 
20 ns of the confinement step. So, with available computer power this method will be easy to use.

This method can be further used for calculating free energy due to very large conformational change. Another application may be
for the case of ligand binding. It can also be used to distinguish the most stable structure during protein design. Another, important
application may be to study the protein folding kinetics, specially to study the phi value analysis.

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
\end{center}
\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
\label{fig:method}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{bhp.pdf}
\end{center}
\caption{Two conformations from $\beta$ hairpin from protein G, bhp1 and bhp2. The two stranded $\beta$ sheet, bhp1 is the native structure and the three stranded $\beta$ sheet is known as bhp3.}
\label{fig:bhp_conf}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
\end{center}
\caption{Two protein with 95 \% similar sequence but different folds}
\label{fig:G95}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.8 in,height=3.0 in]{T0559.pdf}
\end{center}
\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values 
of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
\label{fig:T0559}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.8 in,height=3.0 in]{T0538.pdf}
\end{center}
\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
from the group PconsR, Shell and FOLDIT respectively.}
\label{fig:T0538}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{T0560.pdf}
\end{center}
\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
were from the group "Splicer".}
\label{fig:T0560}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{T0540.pdf}
\end{center}
\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540). 
Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
\label{fig:T0540}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=5.7 in]{T0531.pdf}
\end{center}
\caption{The native structure and 5 models of extracellular domain of the jumping translocation
breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
\label{fig:T0531}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=12cm,height=10cm]{T0605.pdf}
\end{center}
\caption{ Three models of the leucine zipper domain of cGMP dependent protein
kinase I (original pdb id: 3NMD and CASP target: T0605). The original target sequence of this protein was 72 residues.
However only residues 3 to 66 is only kept during result announcement as that part can be found from crystallographic analysis. The
region 3 to 66 are almost identical in all the 3 models  }
\label{fig:T0605}
\end{figure}
